---
title: "Prediction Assignment Writeup"
author: "Yashar F."
date: "October 13, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Motivation

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, we will be using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

## Getting and Cleaning Data
The data for this project come from this source: <http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har>
The following steps were taken to create the tidy data file:
1. Set my working directory 
2. Download the training data from the following link <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>, and save it as “pml-training.csv” file into my working directory.
3. Download the testing data from the following link <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv>, and save it as “pml-testing.csv” file into my working directory.
4. Read in training and testing data.

```{r}
setwd("~/R Projects/Practical Machine Learning/Course Project")
URLTraining <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
URLTesting <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(URLTraining,"./pml-training.csv",mode = "wb") #use mode = "wb" to download as binary
download.file(URLTesting,"./pml-testing.csv",mode = "wb") #use mode = "wb" to download as binary
pml_training <- read.csv(file = "./pml-training.csv") #read in the training data
pml_testing <- read.csv(file = "./pml-testing.csv") #read in the testing data
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
